Do Drivers Dream of Walking? An Investigation of Dissonance in the Primary Mode of Travel from the Perspective of Subjective Wellbeing

### 2019-10-19 Dissonance_Santiago_v1.tex
Commit with first draft of paper

---
title: Do Drivers Dream of Walking? An Investigation of Dissonance in the Primary Mode of Travel from the Perspective of Subjective Wellbeing
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
  Transportation in most of the world has been dominated for decades by a fascination with the automobile. Increasingly, however, it is recognized that to achieve a variety of economic, environmental, and public health policy goals it is important to attract and retain users of public transport and active travel. A challenge with the way transportation policy is developed is that it tends to focus on cost-benefit and utilitarian considerations, which may miss other potential benefits of transportation. For this reason, subjective wellbeing (SWB) has been proposed as a way to enhance our understanding of the preferences and choices of travellers, as well as a way to evaluate the benefits of transportation beyond utilitarian considerations. The objective of this paper is to analyze the modes that people commonly use, and to what extent they are aligned (or not) with a variety of affective and utilitarian values. In other words, we are interested in the potential for dissonance with respect to the primary mode of travel, from the perspective of affective and utilitarian values. The study is based on data collected from a sample of travellers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name the mode or modes that they associated to affective values (freedom, enjoyment, and happiness) and to utilitarian values (safety, time waste, and efficiency). The results indicate that users of public transportation experience the most dissonance in terms of affective values, and active travellers the least. For those travellers who experience dissonance, active travel is the mode most commonly associated with affective values.

  To properly assess transportation policy This poses a number of challenges Broadening the commonly used transport project assessment will be relevant for achieving this goal. While satisfaction with travel has been mainly used in the context of daily trips and mainly under cost-benefit and utilitarian considerations, subjective wellbeing (SWB) over time has risen as an alternative for understanding the preferences and choices of transport users. The aim of this paper is to analyse the factors that can potentially affect the levels of SWB in the wide spectrum of public, private and active transport modes in the context of Santiago de Chile. The paper aims to reflect on the possibilities of increasing the SWB with a view to enhance the attractiveness of public transport and active travel. The paper investigates the affective and instrumental/utilitarian factors of SWB, according to modal choice. It uses an analysis of correspondence that considers the actual primary mode used, versus the mode that each individual associated to both utilitarian and affective SWB factors. The results show that there is more persistent modal dissonance between positive emotional and utilitarian associations within the more disadvantaged groups – particularly among users of public transport. The study shows that apart from time-savings and efficiency, security has also been cross-sectionally sensitive according to the data analysed. Factors such as happiness, enjoyment and freedom require far more attention in transport planning as they are strongly related to the dimensions of subjective and personal wellbeing. Targeting these factors will be relevant for attracting new users to public transportation and active travel.

journal: "An Elsevier journal"
date: "`r Sys.Date()`"
bibliography: mybibfile.bib
#linenumbers: true
#numbersections: true
csl: elsevier-harvard.csl
output: rticles::elsevier_article
---

<!-- This paper is an R markdown document that includes self-contained, reproducible analysis. To reproduce the paper the following files are needed:

1. Dissonance_Santiago_v1.Rmd (this R markdown document)

2. `Dissonance_Santiago.RData`. For reference: this data file was created using R notebook `00. Analysis of Dissonance Detailed.Rmd`. The initial analysis was done in the same R notebook. Neither of these two notebooks are needed for this document.

3. For compiling into a pdf in the style of an Elsevier document: elsarticle.cls, elsevier-harvard.csl, elsevier-harvard-without-titles.csl, elsevier-vancouver.csl, elsevier-without-titles.cls, elsevier-with-titles.csl, elsevier-with-titles-alphabetical.csl, mybibfile.bib
-->

```{r load-packages, include=FALSE}

# Load packages used in this document

library(rticles)
library(readxl)
library(tidyverse)
library(rgdal)
library(sf)
library(ggmosaic)
library(ggthemes)
library(kableExtra)
```

```{r load-data, include=FALSE}

# To run this notebook, the file `Dissonance_Santiago.RData` is needed. This file contains the following object:
# 
# master: a dataframe with information collected from 451 respondents. The variables in the dataframe are:
# 
# - NUMERO: Respondent ID
# - USED: Main mode of transportation used by the respondent
# - FREEDOM: Mode(s) that the respondent associates with feelings of freedom
# - ENJOYMENT: Mode(s) that the respondent associates with feelings of enjoyment
# - HAPPINESS: Mode(s) that the respondent associates with feelings of happiness
# - POVERTY: Mode(s) that the respondent associates with feelings of poverty
# - LUXURY: Mode(s) that the respondent associates with feelings of luxury
# - STATUS: Mode(s) that the respondent associates with feelings of status
# - SAFETY: Mode(s) that the respondent associates with feelings of safety
# - TIME: Mode(s) that the respondent associates with feelings of WASTED TIME
# - EFFICIENCY: Mode(s) that the respondent associates with feelings of efficiency
# - AGE: Age of the respondent
# - EDUCATION: Level of education of the respondent
# - INCOME: Level of income of respondent
# - TRAVEL_TIME: Trip duration of regular commuting trip 
# - WEIGHT_*: A weight for the responses by ATTRIBUTE. Some respondents selected more than one mode that they associated with feelings of {FREEDOM, ENJOYMENT, HAPPINESS, SAFETY, TIME, and EFFICIENCY}. To avoid giving greater weight to respondents who indicated several modes of transportation, a weight was calculated that is the inverse of the number of modes indicated. These weights can be used to weight the frequencies prior to any analysis.
# 
# The modes of transportation (USED, FREEDOM, ENJOYMENT, HAPPINESS, SAFETY, TIME, EFFICIENCY) are coded as: 
# "A": Auto
# "B": Taxi
# "C": Colectivo
# "D": Moto
# "E": Metro
# "F": Bus
# "G": Bicycle
# "H": Walk
#
# AGE is coded as:
#
# "A": Less18
# "B": 18to24
# "C": 25to34
# "D": 35to54
# "E": 55to64
# "F": 65more
# 
# EDUCATION is coded as:
# 
# "A": ELEMENTARY
# "B": SECUNDARY
# "C": PROFESS_TECH
# "D": COLLEGE
# "E": POSTGRAD
#
# INCOME is coded as (in pesos?):
# 
# "A": Less423
# "B": 423to639
# "C": 639to977
# "D": 977to1550
# "E": 1550to2380
# "F": 2380more
# 
# TRAVEL_TIME is coded as:
#
# "A": 0-20 min
# "B": 20-40 min
# "C": 40-60 min
# "D": 1h and more

# Read the RData file:

load("Dissonance_Santiago.RData")
```

```{r data-preparation, include=FALSE}

# The master dataframe is processed for analysis. We plan to conduct two distinct types of analysis:
#
# 1. Analysis of dissonace using the characteristics of respondents (i.e., AGE, EDUCATION, INCOME, (Mode) USED, and TRAVEL_TIME) by ATTRIBUTE (i.e., FREEDOM et al.). In this case, dissonance is defined as follows:
#
# >> If the mode used is identical to the mode that the respondent associates with ATTRIBUTE (i.e. FREEDOM et al.), we consider that there is no dissonance; otherwise, we consider that there is dissonance. For example, if the respondent's main mode of transportation is Auto, and the respondent associates Auto with feelings of FREEDOM, then there is no dissonance. On the other hand, if the respondent's main mode of transportation is Auto, but associates Walking with feelings of FREEDOM, there is dissonance.
# 
# Given this, we will calculate how frequently there is dissonance at different levels of AGE, EDUCATION, INCOME, MODE, TRAVEL_TIME. Tests of independence can be implemented.
#
# 2. Analysis of dissonance using the mode USED and the mode respondents associated with ATTRIBUTE, then stratified by characteristics of respondents (i.e., AGE et al.) in this case, we see how frequently each mode was associated with ATTRIBUTE, given the main mode of transportation of the respondents.
#
# Here the dataframes are prepared to support these two types of analysis

#################################################################################
# Prepare a dataframe for the analysis of dissonance by attribute (Analysis 1.) #
#################################################################################

# First, select the variables needed from the master table, and code dissonance as 1 (no dissonance) or two (dissonance):

Overall_dissonance <- master %>%
  transmute(USED, 
            FREEDOM_DISSONANCE = ifelse(USED == FREEDOM, 1, 2),
            ENJOYMENT_DISSONANCE = ifelse(USED == ENJOYMENT, 1, 2),
            HAPPINESS_DISSONANCE = ifelse(USED == HAPPINESS, 1, 2),
            POVERTY_DISSONANCE = ifelse(USED == POVERTY, 1, 2),
            LUXURY_DISSONANCE = ifelse(USED == LUXURY, 1, 2),
            STATUS_DISSONANCE = ifelse(USED == STATUS, 1, 2),
            SAFETY_DISSONANCE = ifelse(USED == SAFETY, 1, 2),
            TIME_DISSONANCE = ifelse(USED == TIME, 1, 2),
            EFFICIENCY_DISSONANCE = ifelse(USED == EFFICIENCY, 1, 2),
            AGE,
            EDUCATION,
            INCOME,
            TRAVEL_TIME)

# Recode variables and convert to factors:

Overall_dissonance <- Overall_dissonance %>%
  mutate(USED = case_when(USED == "A" ~ "Auto",
                                  USED == "B" | USED == "C" | USED == "D" ~ "Other",
                                  USED == "E" | USED == "F" ~ "Public",
                                  USED == "G" | USED == "H" ~ "Active"),
         AGE = case_when(AGE == "A" | AGE == "B" | AGE == "C" ~ "< 35",
                         AGE == "D" ~ "35 - 54",
                         AGE == "E" | AGE == "F" ~ "> 54"),
         EDUCATION = case_when(EDUCATION == "A" | EDUCATION == "B" ~ "K-12",
                               EDUCATION == "C" | EDUCATION == "D" ~ "Tech/Univ",
                               EDUCATION == "E" ~ "Graduate"),
         INCOME = case_when(INCOME == "A" | INCOME == "B" ~ "Low",
                            INCOME == "C" | INCOME == "D" ~ "Middle",
                            INCOME == "E" | INCOME == "F" ~ "High"),
         TRAVEL_TIME = case_when(TRAVEL_TIME == "A" ~ "< 20 min",
                                 TRAVEL_TIME == "B" ~ "20 - 40 min",
                                 TRAVEL_TIME == "C" ~ "40 - 60 min",
                                 TRAVEL_TIME == "D" ~ "> 60 min")) %>%
  mutate(USED = factor(USED, levels = c("Auto", "Active", "Public", "Other"), labels = c("Auto", "Active", "Public", "Other")),
         FREEDOM_DISSONANCE = factor(FREEDOM_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         ENJOYMENT_DISSONANCE = factor(ENJOYMENT_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         HAPPINESS_DISSONANCE = factor(HAPPINESS_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         POVERTY_DISSONANCE = factor(POVERTY_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         LUXURY_DISSONANCE = factor(LUXURY_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         STATUS_DISSONANCE = factor(STATUS_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         SAFETY_DISSONANCE = factor(SAFETY_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         TIME_DISSONANCE = factor(TIME_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         EFFICIENCY_DISSONANCE = factor(EFFICIENCY_DISSONANCE, levels = c(1, 2), labels = c("No Dissonance", "Dissonance")),
         AGE = factor(AGE, levels = c("< 35", "35 - 54", "> 54"), labels = c("< 35", "35 - 54", "> 54")),
         EDUCATION = factor(EDUCATION, levels = c("K-12", "Tech/Univ", "Graduate"), labels = c("K-12", "Tech/Univ", "Graduate")),
         INCOME = factor(INCOME, levels = c("Low", "Middle", "High"), labels = c("Low", "Middle", "High")),
         TRAVEL_TIME = factor(TRAVEL_TIME, levels = c("< 20 min", "20 - 40 min", "40 - 60 min", "> 60 min"), labels = c("< 20 min", "20 - 40 min", "40 - 60 min", "> 60 min")))

#############################################################################
# Prepare a dataframe for the analysis of dissonance by modes (Analysis 2.) #
#############################################################################

# First, split the responses regarding the modes associated with each attribute, so that we can record each mode selected individually:

Modes_by_Attribute <- master %>% 
    mutate(FREEDOM = strsplit(as.character(FREEDOM), ""),
           ENJOYMENT = strsplit(as.character(ENJOYMENT), ""),
           HAPPINESS = strsplit(as.character(HAPPINESS), ""),
           POVERTY = strsplit(as.character(POVERTY), ""),
           LUXURY = strsplit(as.character(LUXURY), ""),
           STATUS = strsplit(as.character(STATUS), ""),
           SAFETY = strsplit(as.character(SAFETY), ""),
           TIME = strsplit(as.character(TIME), ""),
           EFFICIENCY = strsplit(as.character(EFFICIENCY), ""))

# Here, we convert each of the responses into an individual row and stack the resulting dataframes, after adding the name of the attribute for identification:

Modes_by_Attribute <- rbind(data.frame(Modes_by_Attribute %>% 
                                         select(-c(ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY, 
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY, 
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>% 
                                         rename(IDEAL = FREEDOM, WEIGHT = WEIGHT_FREEDOM) %>%
                                         unnest(IDEAL),
                                       ATTRIBUTE = "FREEDOM"),
                            data.frame(Modes_by_Attribute %>%
                                         select(-c(FREEDOM, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY, 
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY, 
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>% 
                                         rename(IDEAL = ENJOYMENT, WEIGHT = WEIGHT_ENJOYMENT) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "ENJOYMENT"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY, 
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM,
                                                   WEIGHT_ENJOYMENT,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY, 
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>% 
                                         rename(IDEAL = HAPPINESS, WEIGHT = WEIGHT_HAPPINESS) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "HAPPINESS"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY,
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY,
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>%
                                         rename(IDEAL = POVERTY, WEIGHT = WEIGHT_POVERTY) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "POVERTY"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   #LUXURY,
                                                   STATUS,
                                                   SAFETY,
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   #WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY,
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>%
                                         rename(IDEAL = LUXURY, WEIGHT = WEIGHT_LUXURY) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "LUXURY"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   #STATUS,
                                                   SAFETY,
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   #WEIGHT_STATUS,
                                                   WEIGHT_SAFETY,
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>%
                                         rename(IDEAL = STATUS, WEIGHT = WEIGHT_STATUS) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "STATUS"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   TIME, 
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_TIME, 
                                                   WEIGHT_EFFICIENCY)) %>%
                                         rename(IDEAL = SAFETY, WEIGHT = WEIGHT_SAFETY) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "SAFETY"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY,
                                                   EFFICIENCY,
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY,
                                                   WEIGHT_EFFICIENCY)) %>%
                                         rename(IDEAL = TIME, WEIGHT = WEIGHT_TIME) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "TIME"),
                            data.frame(Modes_by_Attribute %>% 
                                         select(-c(FREEDOM, 
                                                   ENJOYMENT, 
                                                   HAPPINESS,
                                                   POVERTY,
                                                   LUXURY,
                                                   STATUS,
                                                   SAFETY,
                                                   TIME, 
                                                   WEIGHT_FREEDOM, 
                                                   WEIGHT_ENJOYMENT, 
                                                   WEIGHT_HAPPINESS,
                                                   WEIGHT_POVERTY,
                                                   WEIGHT_LUXURY,
                                                   WEIGHT_STATUS,
                                                   WEIGHT_SAFETY,
                                                   WEIGHT_TIME)) %>%
                                         rename(IDEAL = EFFICIENCY, WEIGHT = WEIGHT_EFFICIENCY) %>%
                                         unnest(IDEAL), 
                                       ATTRIBUTE = "EFFICIENCY"))

# There are some missing values. Drop all NAs:

Modes_by_Attribute <- drop_na(Modes_by_Attribute)

# We now recode the variables:

Modes_by_Attribute <- Modes_by_Attribute %>%
  mutate(USED = case_when(USED == "A" ~ "Auto",
                                  USED == "B" | USED == "C" | USED == "D" ~ "Other",
                                  USED == "E" | USED == "F" ~ "Public",
                                  USED == "G" | USED == "H" ~ "Active"),
         IDEAL = case_when(IDEAL == "A" ~ "Auto",
                                  IDEAL == "B" | IDEAL == "C" | IDEAL == "D" ~ "Other",
                                  IDEAL == "E" | IDEAL == "F" ~ "Public",
                                  IDEAL == "G" | IDEAL == "H" ~ "Active"),
         AGE = case_when(AGE == "A" | AGE == "B" | AGE == "C" ~ "< 35",
                         AGE == "D" ~ "35 - 54",
                         AGE == "E" | AGE == "F" ~ "> 54"),
         EDUCATION = case_when(EDUCATION == "A" | EDUCATION == "B" ~ "K-12",
                               EDUCATION == "C" | EDUCATION == "D" ~ "Tech/Univ",
                               EDUCATION == "E" ~ "Graduate"),
         INCOME = case_when(INCOME == "A" | INCOME == "B" ~ "Low",
                            INCOME == "C" | INCOME == "D" ~ "Middle",
                            INCOME == "E" | INCOME == "F" ~ "High"),
         TRAVEL_TIME = case_when(TRAVEL_TIME == "A" ~ "< 20 min",
                                 TRAVEL_TIME == "B" ~ "20 - 40 min",
                                 TRAVEL_TIME == "C" ~ "40 - 60 min",
                                 TRAVEL_TIME == "D" ~ "> 60 min"))

# After recoding the , there will be some duplicate entries, because a person may have attributed, say, feelings of freedom to walking and cycling, so in the table this will appear as two instances of "Active". This is ok, because each is weighted appropriately.

# Convert to factors:

Modes_by_Attribute <- Modes_by_Attribute %>%
  mutate(USED = factor(USED, levels = c("Auto", "Active", "Public", "Other"), labels = c("Auto", "Active", "Public", "Other")),
         IDEAL = factor(IDEAL, levels = c("Auto", "Active", "Public", "Other"), labels = c("Auto", "Active", "Public", "Other")),
         AGE = factor(AGE, levels = c("< 35", "35 - 54", "> 54"), labels = c("< 35", "35 - 54", "> 54")),
         EDUCATION = factor(EDUCATION, levels = c("K-12", "Tech/Univ", "Graduate"), labels = c("K-12", "Tech/Univ", "Graduate")),
         INCOME = factor(INCOME, levels = c("Low", "Middle", "High"), labels = c("Low", "Middle", "High")),
         TRAVEL_TIME = factor(TRAVEL_TIME, levels = c("< 20 min", "20 - 40 min", "40 - 60 min", "> 60 min"), labels = c("< 20 min", "20 - 40 min", "40 - 60 min", "> 60 min")))
```

```{r cross-tabs-function, include=FALSE}

## Define functions for crosstabs; these functions enhance the base crosstabs functions and make it easier to report cross tabulations in different ways.

crosstab <- function (..., dec.places = NULL,
                           type = NULL,
                           style = "wide",
                           row.vars = NULL,
                           col.vars = NULL,
                           percentages = TRUE, 
                           addmargins = TRUE,
                           subtotals=TRUE)

###################################################################################
#                                                                                 #
# Function created by Dr Paul Williamson, Dept. of Geography and Planning,        #
# School of Environmental Sciences, University of Liverpool, UK.                  #
#                                                                                 #
# Adapted from the function ctab() in the catspec packge.                         #
#                                                                                 #
# Version: 12th July 2013                                                         #
#                                                                                 #
# Output best viewed using the companion function print.crosstab()                #
#                                                                                 #
###################################################################################

  #Declare function used to convert frequency counts into relevant type of proportion or percentage
  {mk.pcnt.tbl <- function(tbl, type) {
        a <- length(row.vars)
        b <- length(col.vars)
        mrgn <- switch(type, column.pct = c(row.vars[-a], col.vars), 
                            row.pct = c(row.vars, col.vars[-b]),
                            joint.pct = c(row.vars[-a], col.vars[-b]),
                            total.pct = NULL)
        tbl <- prop.table(tbl, mrgn)
        if (percentages) {
            tbl <- tbl * 100
        }
        tbl
    }

    #Find no. of vars (all; row; col) for use in subsequent code
    n.row.vars <- length(row.vars)
    n.col.vars <- length(col.vars)
    n.vars <- n.row.vars + n.col.vars

    #Check to make sure all user-supplied arguments have valid values
    stopifnot(as.integer(dec.places) == dec.places, dec.places > -1)
    #type: see next section of code
    stopifnot(is.character(style))    
    stopifnot(is.logical(percentages))
    stopifnot(is.logical(addmargins))
    stopifnot(is.logical(subtotals))
    stopifnot(n.vars>=1)

    #Convert supplied table type(s) into full text string (e.g. "f" becomes "frequency")
    #If invalid type supplied, failed match gives user automatic error message
    types <- NULL
    choices <- c("frequency", "row.pct", "column.pct", "joint.pct", "total.pct")
    for (tp in type) types <- c(types, match.arg(tp, choices))
    type <- types

    #If no type supplied, default to 'frequency + total' for univariate tables and to
    #'frequency' for multi-dimenstional tables

>>>>

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{An Elsevier journal} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Do Drivers Dream of Walking? An Investigation of Dissonance in the Primary Mode of Travel from the Perspective of Subjective Wellbeing},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}
\begin{frontmatter}

  \title{Do Drivers Dream of Walking? An Investigation of Dissonance in the
Primary Mode of Travel from the Perspective of Subjective Wellbeing}
    \author[Some University]{Author 1\corref{c1}}
   \ead{a1@university.com} 
   \cortext[c1]{Corresponding Author}
    \author[Some School]{Author 2}
   \ead{a2@school.com}

      \address[Some University]{Department, Street, City, State, Zip}
    \address[Some School]{Department, Street, City, State, Zip}

  \begin{abstract}
  Transportation in most of the world has been dominated for decades by a
  fascination with the automobile. Increasingly, however, it is recognized
  that to achieve a variety of economic, environmental, and public health
  policy goals it is important to attract and retain users of public
  transport and active travel. A challenge with the way transportation
  policy is developed is that it tends to focus on cost-benefit and
  utilitarian considerations, which may miss other potential benefits of
  transportation. For this reason, subjective wellbeing (SWB) has been
  proposed as a way to enhance our understanding of the preferences and
  choices of travellers, as well as a way to evaluate the benefits of
  transportation beyond utilitarian considerations. The objective of this
  paper is to analyze the modes that people commonly use, and to what
  extent they are aligned (or not) with a variety of affective and
  utilitarian values. In other words, we are interested in the potential
  for dissonance with respect to the primary mode of travel, from the
  perspective of affective and utilitarian values. The study is based on
  data collected from a sample of travellers in the city of Santiago, in
  Chile. Participants in the study were asked about their usual mode of
  travel, and then were asked to name the mode or modes that they
  associated to affective values (freedom, enjoyment, and happiness) and
  to utilitarian values (safety, time waste, and efficiency). The results
  indicate that users of public transportation experience the most
  dissonance in terms of affective values, and active travellers the
  least. For those travellers who experience dissonance, active travel is
  the mode most commonly associated with affective values.

  To properly assess transportation policy This poses a number of
  challenges Broadening the commonly used transport project assessment
  will be relevant for achieving this goal. While satisfaction with travel
  has been mainly used in the context of daily trips and mainly under
  cost-benefit and utilitarian considerations, subjective wellbeing (SWB)
  over time has risen as an alternative for understanding the preferences
  and choices of transport users. The aim of this paper is to analyse the
  factors that can potentially affect the levels of SWB in the wide
  spectrum of public, private and active transport modes in the context of
  Santiago de Chile. The paper aims to reflect on the possibilities of
  increasing the SWB with a view to enhance the attractiveness of public
  transport and active travel. The paper investigates the affective and
  instrumental/utilitarian factors of SWB, according to modal choice. It
  uses an analysis of correspondence that considers the actual primary
  mode used, versus the mode that each individual associated to both
  utilitarian and affective SWB factors. The results show that there is
  more persistent modal dissonance between positive emotional and
  utilitarian associations within the more disadvantaged groups --
  particularly among users of public transport. The study shows that apart
  from time-savings and efficiency, security has also been
  cross-sectionally sensitive according to the data analysed. Factors such
  as happiness, enjoyment and freedom require far more attention in
  transport planning as they are strongly related to the dimensions of
  subjective and personal wellbeing. Targeting these factors will be
  relevant for attracting new users to public transportation and active
  travel.\\
  \end{abstract}

 \end{frontmatter}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Transportation planning for decades has focused on providing mobility
for the private automobile. This is a model of development that was
initially introduced in North America as a solution to urbanizations
problems, and that was eventually copied elsewhere (Brown et al., 2009;
Angotti, 1996) . Despite the initial promise of automotive technology,
it is now evident that mobility centred on the private automobile has
given rise to a litany of maladies that are in urgent need of
correction. This includes environmental concerns (i.e., climate change;
Chapman, 2007) as well as other social (Boschmann and Kwan, 2008),
health (Khreis et al., 2016), and equity issues (Bocarejo and Oviedo,
2012). As the real impacts of our societal dependence on the private
autombobile have become increasingly evident, the transportation agenda
has in recent years aimed to shift focus to the reduction of car-use and
towards the creation of mobility polycultures that offer a broader menu
of transportation alternatives than just the private automobile (Miller,
2011) . In order to successfully achieve this goal, it is essential not
only to provide the services and facilities that support public
transportation and active travel, but also to attract new users to these
modes of transportation (Ettema et al.~2011). Within this context, it
has been argued that sustainable transportation policies require all
participants in the transportation system to challenge what Gossling and
Cohen (2014) term ``transportation taboos'': deep-seated ideas
concerning the contribution to emissions by individuals, the inequality
of market-based approaches, and the social and psychological functions
of transportation. With respect to the latter, moving beyond a purely
utilitarian focus, it is important to understand how users perceive
affective and instrumental elements on their transportation experience,
since this can help to incorporate adequate improvements in alternative
transportation, including transit and active modes, to make them more
attractive to users (Anable and Gatersleben, 2005; Domarchi et al.,
2008; Gatersleben and Uzzell, 2007; Steg 2005 ).

A useful lens to investigate the social and psychological functions of
transportation is by means of the concept of subjective wellbeing (SWB.
See, inter alia, De Vos et al., 2013 and Chatterjee et al., 2019) . SWB
is defined by the OECD as ``{[}g{]}ood mental states, including all of
the various evaluations, positive and negative, that people make of
their lives, and the affective reactions of people to their
experiences'' (OECD, 2013, p.~10). Accordingly, the aim of this paper is
to analyse some of the factors that can potentially affect the levels of
SWB in the wide spectrum of public, private and active transport modes
in the context of Santiago de Chile. Previous research suggests that
public transport users experience a negative gap in desired and actual
travel time, whereas active travellers don't seem to mind somewhat
longer trips than what strictly needed (e.g., Páez and Whalen, 2010). In
the context of a Latin American country, historically the poor travel
experience of public transit users and active travellers reflects the
social inequalities of social groups that have no alternative but to
move in these types of modes while living far from the work centres and
main activities. The paper aims to reflect on the possibilities of
increasing the SWB while travelling, and how this could increase the
attractiveness of public transport and active travel in the
socio-demographic groups studied. This can help not only to reduce the
use of private transport but also decrease the inequality gap between
these users.
